---
title: "autheme_introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{autheme_introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

This package has 2 aims: 

- Get quickly pretty and consistent graphs (ggplot2) and tables (flextable) with a style that suit Aarhus University theme, including colors and fonts. 

- Get quickly a table or grah display of the least square means (emmeans) and p values in table. Also get model diagnostic for linear mixed models.



## Required packages
```{r, message = FALSE, warning=FALSE}
library(extrafont) # to get your device fonts

library(autheme) 

# packages only for the vignette

library(dplyr)  # to summarize tables

```

## Install fonts 

Before starting, install AU fonts on you device. The fonts can be found here:
https://medarbejdere.au.dk/en/administration/communication/guidelines/guidelinesforfonts/downloadfonts/browse


Then import the fonts on R, this has to be done only once.
```{r}
# install.packages('extrafont')
# library(extrafont)
# font_import()
```


```{r, include=FALSE}
windowsFonts("AU Passata" = windowsFont("AU Passata"))
windowsFonts("AU Passata Light" = windowsFont("AU Passata Light"))
```



## Install autheme

Directly from Github:
```{r}
# devtools::install_github("giuliogiagnoni/autheme")
```

Alternatively, download the folder and install: 
```{r}
# install.packages("file_path/file_name.file_extension")
```

# Colors

Palettes from https://medarbejdere.au.dk/en/administration/communication/guidelines/guidelinesforcolours/browse


## All colors (full)
```{r, fig.width = 9, fig.height=5}
show_col(au_pal[["full"]])
```

:::: {style="display: flex;"}
::: {}
## Light colors (light)
```{r}
show_col(au_pal[["light"]])
```
:::
::: {}
## Dark colors (dark)
```{r}
show_col(au_pal[["dark"]])
```
:::
::::

:::: {style="display: flex;"}
::: {}
## Custom set 1: 4 colors (customset1.4)
```{r}
show_col(au_pal[["customset1.4"]])
```
:::
::: {}
## Custom set 2: 5 colors (customset2.5)
```{r}
show_col(au_pal[["customset2.5"]])
```
:::
::::

:::: {style="display: flex;"}
::: {}
## Custom set 3: 6 colors (customset3.6)
```{r}
show_col(au_pal[["customset3.6"]])
```
:::
::: {}
## Custom set 4: 6 colors (customset4.6)
```{r}
show_col(au_pal[["customset4.6"]])
```
:::
::::

## Office themes

11 AU theme for office are present in the package, only 1 color out of 7 changes between the theme but they can be named as follow: "au_pal_blue", "au_pal_purple", ... , "au_pal_black", "au_pal_rainbow"

## Fading colors
If you wan to fade a color, or a set of colors, this can be easily achieve by using the function alpha from the scales package.
```{r}
# library(scales)
# alpha(color, alpha = 0-1)
```


# Scales
The palette can be use with the scale_color_au or scale_fill_au by providing the argumennts palette, as in the example:

```{r}
# scale_color_au(palette = "full)
```

Is possible to fade the color by adding the argument alpha (0-1), as in the example:
```{r}
# scale_color_au(palette = "full, alpha = 0.5)
```

Is also possible to create the scale through scale_color_manual or scale_fill manual, by providing the vector of colors (au_pal_full, au_pal_light, au_pal_dark, ...), as in these two examples:
```{r}
# scale_color_manual(values = au_pal_full) or 
# scale_color_manual(values = au_pal[["full"]])
```



# Themes
The themes available are the same as the standard ggplot theme, with some modification to fit AU style.
There are two version for each theme, one with black and one AU blue text and lines. 

:::: {style="display: flex;"}
::: {}
### theme_au_classic()
```{r, fig.width = 4.5}
ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) +
  geom_point(size = 2) +
 scale_color_au(palette = "customset1.4") +
  theme_au_classic()
```
:::
::: {}
### theme_au_classic_col()
```{r, fig.width = 4.5}
ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) +
  geom_point(size = 2) +
  scale_color_au(palette = "customset1.4") +
  theme_au_classic_col()
```
:::
::::

:::: {style="display: flex;"}
::: {}
### theme_au_bw()
```{r, fig.width = 4.5}
ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) +
  geom_point(size = 2) +
  scale_color_au(palette = "customset1.4") +
  theme_au_bw()
```
:::
::: {}
### theme_au_bw_col()
```{r, fig.width = 4.5}
ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) +
  geom_point(size = 2) +
  scale_color_au(palette = "customset1.4") +
  theme_au_bw_col()
```
:::
::::

## Boxplots

:::: {style="display: flex;"}
::: {}
### theme_au_minimal()
```{r, fig.width = 4.5}
ggplot(iris, aes(Species, Sepal.Width, fill = Species)) +
  geom_boxplot() +
  scale_fill_au(palette = "customset1.4") +
  theme_au_minimal()
```
:::
::: {}
### theme_au_minimal_col()
```{r, fig.width = 4.5}
ggplot(iris, aes(Species, Sepal.Width, fill = Species)) +
  geom_boxplot() +
  scale_fill_au(palette = "customset1.4") +
  theme_au_minimal_col()
```
:::
::::


## Facets

:::: {style="display: flex;"}
::: {}
### theme_AU()
```{r}
ggplot(iris, aes(Sepal.Length, Sepal.Width)) +
  geom_point(size = 2) +
  facet_wrap(~ Species) +
  theme_au_void()
```
:::
::: {}
### theme_AU_color()
```{r}
ggplot(iris, aes(Sepal.Length, Sepal.Width)) +
  geom_point(size = 2) +
  facet_wrap(~ Species) +
  theme_au_void_col()
```
:::
::::




# Tables


### flextable_au_tidy()
```{r}
flextable_au_tidy(iris %>% group_by(Species) %>%
             summarize(Mean_Sepal.Length = mean(Sepal.Length),
                       SD_Sepal.Length = sd(Sepal.Length)))
```

### flextable_au_color()
```{r}
flextable_au_color(iris %>% group_by(Species) %>%
             summarize(Mean_Sepal.Length = mean(Sepal.Length),
                       SD_Sepal.Length = sd(Sepal.Length)))
```




# Model outputs

I include in the package a dataset publicly available (find information in data-raw/Nakajima2021.R). This data set is just used as an example, none of the statistic showed here should be consider relevant.


## Model diagnostic
```{r}
data <- Nakajima2021_bloodparameters
data <- data %>% dplyr::rename(Day =  "Day after the start of grazing (Day 0)",
                                                       Cow = "Individual number")

# set the order of the independent variables, these were already like this they go in alphabetic order
data$Treatment <- factor(data$Treatment, c("Confinement", "Grazing"))

# set model and independent variable of interest
model <- "Treatment * Day + (1|Cow)"
ind_var <- "Treatment"
  
# just some cleaning of this data set
colnames(data) <-  sub("\r\n", "", colnames(data))
colnames(data) <-  sub(" \r\n", "", colnames(data))
colnames(data) <-  gsub(" ", "", colnames(data))

colnames(data) <-  sub("\\(", "_", colnames(data))
colnames(data) <-  sub("\\)", "", colnames(data))
colnames(data) <-  gsub("/", ".", colnames(data))

dep_var <- names(data)[22:30]
```


```{r, fig.width = 9, fig.height=9}
lmerdiagnostic(data, model, ind_var, dep_var[1])
```

## Emmeans table

Fast table output as data frame. Consider if you get some working to got back to the diagnostic of the model.
```{r}
dep_var <- rev(names(data)[22:30])

mytab <- lmertableloop(data, model, ind_var, dep_var, p = "std1", s = 3)
  
mytab

flextable_au_color(mytab)
```

## Emmeans plot

```{r, fig.width=4.5, fig.height=4.5}
model <- "BUN_mg.dl ~ Treatment * Day + (1|Cow)"

plotmeans1(data, model, ind_var)
```





# Troubleshooting

## Legend
### Unwanted legend
```{r}
# guides(fill/color/shape=FALSE) +
```

### Out of margins
This can be easily fixed by spreading the legend on multiple rows
```{r}
#   guides(color = guide_legend(nrow = 2)) +
```


